---
title: "MPE Manuscript Analysis 2021"
author: "Wesley Hull"
date: "11/7/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r, echo=FALSE}
# Packages

library(lattice)
library(psych)
library(nlme)
library(lme4)

library(Rmisc)
library(cowplot)
library(ggpubr)

library(grid)
library(MASS)
library(blmeco)
library(car)
library(AICcmodavg)
library(fitdistrplus)

library(dplyr)
library(MuMIn)
library(lsmeans)
```

## Data Files Needed in Analysis 

  Recruitment data for R 2020  
  
  Field Density Data for R 2019
  
  Mussel.Survival.Data
  
  Pisaster.Measurements.All.2015
  
  Crab.Measurements.All.2015
  
```{r, echo=FALSE}
Recruitment_data_for_R_2020 <- read_csv("C:/Users/wwhul/OneDrive/Manuscripts/MPE PAPER/Data/Recruitment data for R 2020.csv")
Field_Density_Data_For_R_2019 <- read_csv("C:/Users/wwhul/OneDrive/Manuscripts/MPE PAPER/Data/Field Density Data for R 2019.csv")
Mussel_Survival_Data <- read_csv("C:/Users/wwhul/OneDrive/Manuscripts/MPE PAPER/Data/Mussel.Survival.Data.csv")
Pisaster_Measurements_All_2015 <- read_csv("C:/Users/wwhul/OneDrive/Manuscripts/MPE PAPER/Data/Pisaster.Measurements.All.2015.csv")
Crab_Meausrements_All_2015 <- read_csv("C:/Users/wwhul/OneDrive/Manuscripts/MPE PAPER/Data/Crab.Meausrements.All.2015.csv")
```
# RECRUITMENT ANALYSIS 
 Goals
    1) Determine if mussel recruitment differs across our field sites.
    2) Determine if mussel recruitment differs between horizontal low and vertical high surfaces
```{r, echo = TRUE}
# New column dividing recruit count by days in the field
  Recruitment_data_for_R_2020$Rec.Tuff.Day<-Recruitment_data_for_R_2020$Count/104
  Recruitment_data_for_R_2020$Log.Rec.Day<-log(Recruitment_data_for_R_2020$Rec.Tuff.Day)
  
  head(Recruitment_data_for_R_2020)

# Full model
  Mod1<-lm(Recruitment_data_for_R_2020$Log.Rec.Day~Recruitment_data_for_R_2020$Orientation*Recruitment_data_for_R_2020$Site)
  summary(Mod1)

# ANOVA
  Mod1.aov<-aov(Mod1)
  summary(Mod1.aov)     # Interaction is not significant, so remove it and re-run model

# New model with/out interaction
  Mod2<-lm(Recruitment_data_for_R_2020$Log.Rec.Day~Recruitment_data_for_R_2020$Orientation+Recruitment_data_for_R_2020$Site)
  summary(Mod2)

# New ANOVA
  Mod2.aov<-aov(Mod2)
  summary(Mod2.aov)     #Significant effect of site and orientation
                        #Run Tukeys to determine differences between sites

# Tukeys
  TukeyHSD(Mod2.aov)
        # Findings
          # Overall significant difference in recruitment between vertical and horizontal surfaces
          # Significantly more recruitment at Point St. George than Devil's Gate and Belinda Point
          # No difference in recruitment between Devil's Gate and Belinda Point
```

# DENSITY ANALYSIS 
 Goals:
    1) Determine if predator density differs across our field sites.
    2) Determine if predator density differs at each site
```{r, echo=TRUE}
# New column with log transformed Density data 
  Field_Density_Data_For_R_2019$Log.Density<-log(Field_Density_Data_For_R_2019$Density)
  head(Field_Density_Data_For_R_2019)
  
# Full Model
  Density.lm<-lm(Field_Density_Data_For_R_2019$Log.Density~Field_Density_Data_For_R_2019$Predator*Field_Density_Data_For_R_2019$Site)
  summary(Density.lm)

# ANOVA
  Density.aov<-aov(Density.lm)
  summary(Density.aov)      # Interaction is not significant, re-run without interaction

# New model with/out interaction
  Density.lm2<-lm(Field_Density_Data_For_R_2019$Log.Density~Field_Density_Data_For_R_2019$Predator+Field_Density_Data_For_R_2019$Site)
  summary(Density.lm2)      

# New ANOVA
  Density.aov2<-aov(Density.lm2)
  summary(Density.aov2)     # Significant effect of predator type and site

# Tukey's
  TukeyHSD(Density.aov2)    
          # Findings
              # Overall significant difference in seastar and crab abundance
              # Point St. George Had the highest predator density followed by Devil's Gate and Belinda Point

```
  
# MUSSEL SURVIVAL ANALYSES 
### Proportion of Living 
```{r, echo=TRUE}
# Random affect is nested: blocks are repeated within sites, so code should be (1|Block:Site) or (1|Site:Block).
    
# With random intercept
    Mod1<-lmer(Proportion.Living~Treatment.S*Treatment.C*Site+(1|Site:Block), data = Mussel_Survival_Data)
    summary(Mod1)       #Variation from block is extremely low
    
# Without random intercept
    Mod2<-lm(Proportion.Living~Treatment.S*Treatment.C*Site,data = Mussel_Survival_Data)
    summary(Mod2)
      
# Likelihood ratio test
    anova(Mod1,Mod2)    # No difference between models, 
                        # Drop effect of block from model
    
# Model Selection
    
  # full
    Full.lm<-lm(Proportion.Living~Treatment.S*Treatment.C*Site, data = Mussel_Survival_Data)
      
# model comparison function
    options(na.action=na.fail)
    
    dredge(Full.lm)   # Full model is the best model
      
# Model summary/anova
    summary(Full.lm)
    anova(Full.lm)      # Run pairwise comparisons
    
# Pairwise comparisons
    lsmeans(Full.lm, pairwise ~ Site*Treatment.C*Treatment.S, adjust = "tukey")
  
```
### Natural Log Mussel Survival per Day 
```{r, echo=TRUE}
# Random effect assessment
# Random affect is nested: blocks are repeated within sites, so code should be (1|Block:Site) or (1|Site:Block).
      
# With random intercept
    Mod3<-lmer(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site+(1|Site:Block), data = Mussel_Survival_Data)
    summary(Mod3)       # Seems that the variation from block is extremely low
      
# Without random intercept
    Mod4<-lm(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site,data = Mussel_Survival_Data)
    summary(Mod4)
      
# Likelihood ratio test
    anova(Mod3,Mod4)    # No difference between models, so drop effect of block from model
      
# Model Selection
  
# full
    Full.LN.lm<-lm(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site, data = Mussel_Survival_Data)
      
# model selection
    options(na.action=na.fail)
      
    dredge(Full.LN.lm)   # Full model is the best model
      
# Model summary/anova
    summary(Full.LN.lm)
    anova(Full.LN.lm)  
    
# Pairwise comparisons
    lsmeans(Full.lm, pairwise ~ Site*Treatment.C*Treatment.S, adjust = "tukey")
```
  
# INTERACTION STRENGTH ANALYSIS 
### Per Population Interaction Strength 
```{r, echo=TRUE}
# Remove control treatments    
  Mussel.Survival<-Mussel_Survival_Data[-c(19,20,21,22,23,24,43,44,45,46,47,48,67,68,69,70,71,72),]

# Full Model
    Full.PerCap.glm<-glm(Mussel.Survival$Per.Cap.Int.Str~Mussel.Survival$Treatments.All*Mussel.Survival$Site, family = gaussian())
    summary(Full.PerCap.glm)
    
# ANOVA
    Full.PerCap.aov<-aov(Full.PerCap.glm)
    summary(Full.PerCap.aov)      # Interaction is significant
    
# Tukey's 
    TukeyHSD(Full.PerCap.aov)

```

### Per Capita Interaction Strength 
```{r, echo=TRUE}    
# Full Model
    Full.Pop.lm<-lm(Mussel.Survival$Pop.Int.Str~Mussel.Survival$Treatments.All*Mussel.Survival$Site)
    summary(Full.Pop.lm)
    
# ANOVA
    Full.Pop.aov<-aov(Full.Pop.lm)
    summary(Full.Pop.aov)      # Interaction is significant
    
# Tukey's 
    TukeyHSD(Full.Pop.aov) 
```
    
# SUPPLEMENTAL ANALYSIS 
```{r, echo=TRUE}
# ANOVA
# Sea star
  Rad<-lm(`Star Radius`~ Site, data = Pisaster_Measurements_All_2015)
  summary(Rad)
  Rad.aov<-aov(Rad)
  summary(Rad.aov)
  
# Crabs
  CP<-lm(`Carapace Width (mm)` ~ Site, data = Crab_Meausrements_All_2015)  
  summary(CP)
  CP.aov<-aov(CP)
  summary(CP.aov)  
  
  par(mfrow = c(2,2))
  plot(Rad.aov)   #violates assumptions, run kruskal wallis
  plot(CP.aov)    #violates assumptions, run kruskal wallis
    
# Kruskal wallis
  NP<-kruskal.test(Pisaster_Measurements_All_2015$Site ,Pisaster_Measurements_All_2015$`Star Radius`)
  NP  
  CP<-kruskal.test(Crab_Meausrements_All_2015$Site,Crab_Meausrements_All_2015$`Carapace Width (mm)`)
  CP
```  
