---
title: "MPE Manuscript Analysis 2021"
author: "Wesley Hull"
date: "11/7/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r, echo=FALSE}
# Packages
library(tidyverse)
library(plyr)
library(Rmisc)
library(ggplot2)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(glmmTMB)
library(DHARMa)
library(MuMIn)
library(lme4)
library(lsmeans)
library(car)
library(cowplot)
```

## Data Files Needed in Analysis 

  Recruitment data for R 2020  
  
  Field Density Data for R 2019
  
  Mussel.Survival.Data
  
  Pisaster.Measurements.All.2015
  
  Crab.Measurements.All.2015
  
```{r, echo=FALSE}
Recruitment_data_for_R_2020 <- read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Recruitment%20data%20for%20R%202020.csv")
Field_Density_Data_For_R_2019 <- read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Field%20Density%20Data%20For%20R%202019.csv")
Mussel_Survival_Data <- read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Mussel.Survival.Data.csv")
Pisaster_Measurements_All_2015 <-read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Pisaster.Measurements.All.2015.csv")
Crab_Meausrements_All_2015 <- read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Crab.Meausrements.All.2015.csv")
```
# RECRUITMENT ANALYSIS 
 Goals
    1) Determine if mussel recruitment differs across our field sites.
    2) Determine if mussel recruitment differs between horizontal low and vertical high surfaces
```{r, echo = TRUE}
# New column dividing recruit count by days in the field
  Recruitment_data_for_R_2020$Rec.Tuff.Day<-Recruitment_data_for_R_2020$Count/104
  Recruitment_data_for_R_2020$Log.Rec.Day<-log(Recruitment_data_for_R_2020$Rec.Tuff.Day)
  
  head(Recruitment_data_for_R_2020)

# Full model
  Mod1<-lm(Recruitment_data_for_R_2020$Log.Rec.Day~Recruitment_data_for_R_2020$Orientation*Recruitment_data_for_R_2020$Site)
  summary(Mod1)

# ANOVA
  Mod1.aov<-aov(Mod1)
  summary(Mod1.aov)     # Interaction is not significant, so remove it and re-run model

# New model with/out interaction
  Mod2<-lm(Recruitment_data_for_R_2020$Log.Rec.Day~Recruitment_data_for_R_2020$Orientation+Recruitment_data_for_R_2020$Site)
  summary(Mod2)

# New ANOVA
  Mod2.aov<-aov(Mod2)
  summary(Mod2.aov)     #Significant effect of site and orientation
                        #Run Tukeys to determine differences between sites

# Tukeys
  TukeyHSD(Mod2.aov)
        # Findings
          # Overall significant difference in recruitment between vertical and horizontal surfaces
          # Significantly more recruitment at Point St. George than Devil's Gate and Belinda Point
          # No difference in recruitment between Devil's Gate and Belinda Point
```

# DENSITY ANALYSIS 
 Goals:
    1) Determine if predator density differs across our field sites.
    2) Determine if predator density differs at each site
```{r, echo=TRUE}
# New column with log transformed Density data 
  Field_Density_Data_For_R_2019$Log.Density<-log(Field_Density_Data_For_R_2019$Density)
  head(Field_Density_Data_For_R_2019)
  
# Full Model
  Density.lm<-lm(Field_Density_Data_For_R_2019$Log.Density~Field_Density_Data_For_R_2019$Predator*Field_Density_Data_For_R_2019$Site)
  summary(Density.lm)

# ANOVA
  Density.aov<-aov(Density.lm)
  summary(Density.aov)      # Interaction is not significant, re-run without interaction

# New model with/out interaction
  Density.lm2<-lm(Field_Density_Data_For_R_2019$Log.Density~Field_Density_Data_For_R_2019$Predator+Field_Density_Data_For_R_2019$Site)
  summary(Density.lm2)      

# New ANOVA
  Density.aov2<-aov(Density.lm2)
  summary(Density.aov2)     # Significant effect of predator type and site

# Tukey's
  TukeyHSD(Density.aov2)    
          # Findings
              # Overall significant difference in seastar and crab abundance
              # Point St. George Had the highest predator density followed by Devil's Gate and Belinda Point

```
  
# MUSSEL SURVIVAL ANALYSES 
### Proportion of Living 
```{r, echo=TRUE}
# Random affect is nested: blocks are repeated within sites, so code should be (1|Block:Site) or (1|Site:Block).
    
# With random intercept
    Mod1<-lmer(Proportion.Living~Treatment.S*Treatment.C*Site+(1|Site:Block), data = Mussel_Survival_Data)
    summary(Mod1)       #Variation from block is extremely low
    
# Without random intercept
    Mod2<-lm(Proportion.Living~Treatment.S*Treatment.C*Site,data = Mussel_Survival_Data)
    summary(Mod2)
      
# Likelihood ratio test
    anova(Mod1,Mod2)    # No difference between models, 
                        # Drop effect of block from model
    
# Model Selection
    
  # full
    Full.lm<-lm(Proportion.Living~Treatment.S*Treatment.C*Site, data = Mussel_Survival_Data)
      
# model comparison function
    options(na.action=na.fail)
    
    dredge(Full.lm)   # Full model is the best model
      
# Model summary/anova
    summary(Full.lm)
    anova(Full.lm)      # Run pairwise comparisons
    
# Pairwise comparisons
    lsmeans(Full.lm, pairwise ~ Site*Treatment.C*Treatment.S, adjust = "tukey")
  
```

### Natural Log Mussel Survival per Day 
```{r, echo=TRUE}
# Random effect assessment
# Random affect is nested: blocks are repeated within sites, so code should be (1|Block:Site) or (1|Site:Block).
      
# With random intercept
    Mod3<-lmer(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site+(1|Site:Block), data = Mussel_Survival_Data)
    summary(Mod3)       # Seems that the variation from block is extremely low
      
# Without random intercept
    Mod4<-lm(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site,data = Mussel_Survival_Data)
    summary(Mod4)
      
# Likelihood ratio test
    anova(Mod3,Mod4)    # No difference between models, so drop effect of block from model
      
# Model Selection
  
# full
    Full.LN.lm<-lm(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site, data = Mussel_Survival_Data)
      
# model selection
    options(na.action=na.fail)
      
    dredge(Full.LN.lm)   # Full model is the best model
      
# Model summary/anova
    summary(Full.LN.lm)
    anova(Full.LN.lm)  
    
# Pairwise comparisons
    lsmeans(Full.lm, pairwise ~ Site*Treatment.C*Treatment.S, adjust = "tukey")
```
  
# INTERACTION STRENGTH ANALYSIS 
### Per Population Interaction Strength 
#### How much does the entire predator population at a site reduce mussel survival?
```{r, echo=TRUE}
# Remove control treatments    
  Mussel.Survival<-Mussel_Survival_Data[-c(19,20,21,22,23,24,43,44,45,46,47,48,67,68,69,70,71,72),]

# Full Model
    Full.PerCap.glm<-glm(Mussel.Survival$Per.Cap.Int.Str~Mussel.Survival$Treatments.All*Mussel.Survival$Site, family = gaussian())
    summary(Full.PerCap.glm)
    
# ANOVA
    Full.PerCap.aov<-aov(Full.PerCap.glm)
    summary(Full.PerCap.aov)      # Interaction is significant
    
# Tukey's 
    TukeyHSD(Full.PerCap.aov)

```

### Per Capita Interaction Strength 
```{r, echo=TRUE}    
# Full Model
    Full.Pop.lm<-lm(Mussel.Survival$Pop.Int.Str~Mussel.Survival$Treatments.All*Mussel.Survival$Site)
    summary(Full.Pop.lm)
    
# ANOVA
    Full.Pop.aov<-aov(Full.Pop.lm)
    summary(Full.Pop.aov)      # Interaction is significant
    
# Tukey's 
    TukeyHSD(Full.Pop.aov) 
```
    
# SUPPLEMENTAL ANALYSIS 
```{r, echo=TRUE}
# ANOVA
# Sea star
  Rad<-lm(`Star Radius`~ Site, data = Pisaster_Measurements_All_2015)
  summary(Rad)
  Rad.aov<-aov(Rad)
  summary(Rad.aov)
  
# Crabs
  CP<-lm(`Carapace Width (mm)` ~ Site, data = Crab_Meausrements_All_2015)  
  summary(CP)
  CP.aov<-aov(CP)
  summary(CP.aov)  
  
  par(mfrow = c(2,2))
  plot(Rad.aov)   #violates assumptions, run kruskal wallis
  plot(CP.aov)    #violates assumptions, run kruskal wallis
    
# Kruskal wallis
  NP<-kruskal.test(Pisaster_Measurements_All_2015$Site ,Pisaster_Measurements_All_2015$`Star Radius`)
  NP  
  CP<-kruskal.test(Crab_Meausrements_All_2015$Site,Crab_Meausrements_All_2015$`Carapace Width (mm)`)
  CP
```  




# model selection, PCIS and PPIS
```{r}
# PPIS = total population level impact of each predator type per site
  # PPIS = PPIS=(effect of predator on survival(effect size))×(predator abundance)
    # effect size  =  survival in the respective absent treatment - the survival in the predator treatment

library(dplyr)
library(emmeans)

# -------------------------------------------------
# 1. FULL FACTORIAL MODEL
# -------------------------------------------------

mod_full <- glmmTMB(
  cbind(Survived, 30 - Survived) ~ Site * Treatment.C * Treatment.S +
    (1 | Site:Block),
  family = binomial,
  data = Mussel_Survival_Data
)

summary(mod_full)


# -------------------------------------------------
# 2. MODEL SIMPLIFICATION
# -------------------------------------------------

mod2 <- update(mod_full, . ~ . - Site:Treatment.C:Treatment.S)
mod3 <- update(mod2, . ~ . - Site:Treatment.S)
mod4 <- update(mod3, . ~ . - Site:Treatment.C)

anova(mod_full, mod2, mod3, mod4)
Anova(mod2, type = "II")


# -------------------------------------------------
# 3. DIAGNOSTICS
# -------------------------------------------------

sim1 <- DHARMa::simulateResiduals(mod2)
plot(sim1)


# -------------------------------------------------
# 4. POST-HOC MARGINAL MEANS (FACTORIAL STRUCTURE)
# -------------------------------------------------

library(emmeans)

emm_df <- emmeans(
  mod2,
  ~ Treatment.C * Treatment.S | Site,
  type = "response"
) %>%
  as.data.frame()

emm_df
# -------------------------------------------------
# 5. EXTRACT TREATMENT COMBINATIONS
# -------------------------------------------------

baseline  <- emm_df %>% dplyr::filter(Treatment.C == "Absent",
                                      Treatment.S == "Absent")

crab_only <- emm_df %>% dplyr::filter(Treatment.C == "Present",
                                      Treatment.S == "Absent")

star_only <- emm_df %>% dplyr::filter(Treatment.C == "Absent",
                                      Treatment.S == "Present")

both      <- emm_df %>% dplyr::filter(Treatment.C == "Present",
                                      Treatment.S == "Present")


# -------------------------------------------------
# 6. EFFECT SIZES (SIMPLE ΔS CONTRASTS)
# -------------------------------------------------

# Crab effect
deltaS_crab <- data.frame(
  Site = baseline$Site,
  delta = baseline$prob - crab_only$prob
)

# Sea star effect
deltaS_star <- data.frame(
  Site = baseline$Site,
  delta = baseline$prob - star_only$prob
)

# Combined treatment effect
deltaS_both <- data.frame(
  Site = baseline$Site,
  delta = baseline$prob - both$prob
)


# -------------------------------------------------
# 7. PREDATOR DENSITY (SITE LEVEL)
# -------------------------------------------------

pred_site <- Field_Density_Data_For_R_2019 %>%
  dplyr::group_by(Site, Predator) %>%
  dplyr::summarise(
    mean_density = mean(Density, na.rm = TRUE),
    sd_density   = sd(Density, na.rm = TRUE),
    n_transects  = dplyr::n(),
    se_density   = sd_density / sqrt(n_transects),
    .groups = "drop"
  )

crab_density <- pred_site %>% dplyr::filter(Predator == "Crab")
star_density <- pred_site %>% dplyr::filter(Predator == "Sea star")

both_density <- Field_Density_Data_For_R_2019 %>%
  dplyr::group_by(Site) %>%
  dplyr::summarise(
    combined_density = sum(Density, na.rm = TRUE),
    .groups = "drop"
  )


# -------------------------------------------------
# 8. MERGE EFFECTS WITH DENSITY
# -------------------------------------------------

ppis_crab <- merge(deltaS_crab, crab_density, by = "Site")
ppis_star <- merge(deltaS_star, star_density, by = "Site")
ppis_both <- merge(deltaS_both, both_density, by = "Site")


# -------------------------------------------------
# 9. COMPUTE PPIS + PCIS (CLEAN DEFINITIONS)
# -------------------------------------------------

ppis_crab <- ppis_crab %>%
  dplyr::mutate(
    PPIS       = delta,
    per_capita = delta / mean_density,
    Predator   = "Crab"
  )

ppis_star <- ppis_star %>%
  dplyr::mutate(
    PPIS       = delta,
    per_capita = delta / mean_density,
    Predator   = "Sea star"
  )

ppis_both <- ppis_both %>%
  dplyr::mutate(
    PPIS       = delta,
    per_capita = delta / combined_density,
    Predator   = "Both"
  )


# -------------------------------------------------
# 10. FINAL COMBINED DATASET
# -------------------------------------------------

ppis_all <- dplyr::bind_rows(
  ppis_crab %>% dplyr::select(Site, Predator, delta, PPIS, per_capita),
  ppis_star %>% dplyr::select(Site, Predator, delta, PPIS, per_capita),
  ppis_both %>% dplyr::select(Site, Predator, delta, PPIS, per_capita)
)









# =================================================
# GOLD-STANDARD PPIS + PCIS PIPELINE
# =================================================

library(dplyr)
library(emmeans)

# -------------------------------------------------
# 1. FITTED MODEL (assumed already run)
# mod2
# -------------------------------------------------


# -------------------------------------------------
# 2. FULL FACTORIAL EMMeans (for structure)
# -------------------------------------------------

emm <- emmeans(mod2,
               ~ Treatment.C * Treatment.S | Site,
               type = "response")

contr <- contrast(
  emm,
  method = list(
    Crab = c(-1, 1, 0, 0),
    Star = c(-1, 0, 1, 0),
    Both = c(-1, 0, 0, 1)
  ),
  by = "Site"
)

ppis_df <- as.data.frame(contr)
# -------------------------------------------------
# 4. MERGE DENSITY DATA
# -------------------------------------------------

crab_density <- crab_density %>%
  rename(density = mean_density)

star_density <- star_density %>%
  rename(density = mean_density)

both_density <- both_density %>%
  rename(density = combined_density)

ppis_df <- ppis_df %>%
  mutate(
    Predator = rep(c("Crab", "Sea star", "Both"), length(unique(Site)))
  ) %>%
  left_join(
    bind_rows(
      crab_density %>% mutate(Predator = "Crab"),
      star_density %>% mutate(Predator = "Sea star"),
      both_density %>% mutate(Predator = "Both")
    ),
    by = c("Site", "Predator")
  )


# -------------------------------------------------
# 5. PCIS + DELTA METHOD UNCERTAINTY
# -------------------------------------------------

ppis_df <- ppis_df %>%
  mutate(
    PCIS = PPIS / density,

    # Delta method SE approximation:
    PCIS_SE = sqrt((PPIS_SE / density)^2)
  )


# -------------------------------------------------
# 6. 95% CONFIDENCE INTERVALS
# -------------------------------------------------

ppis_df <- ppis_df %>%
  mutate(
    PPIS_LCL = PPIS - 1.96 * PPIS_SE,
    PPIS_UCL = PPIS + 1.96 * PPIS_SE,

    PCIS_LCL = PCIS - 1.96 * PCIS_SE,
    PCIS_UCL = PCIS + 1.96 * PCIS_SE
  )


# -------------------------------------------------
# 7. FINAL OUTPUT
# -------------------------------------------------

ppis_all <- ppis_df %>%
  select(Site, Predator,
         PPIS, PPIS_SE, PPIS_LCL, PPIS_UCL,
         PCIS, PCIS_SE, PCIS_LCL, PCIS_UCL)
```



```{r}

# Recode Predator factor
ppis_all <- ppis_all %>%
  mutate(
    Predator = factor(Predator, levels = c("Sea star", "Crab", "Both"))
  )

# PCIS plot with manual colors hardcoded
PCIS<- ggplot(ppis_all, aes(x = per_capita, y = Site, fill = Predator)) +
  geom_hline(yintercept = seq(1,3), linewidth = 1, color = "gray70") +
  geom_point(shape = 21, size = 7, position = position_dodge(0.0)) +
    scale_x_continuous(breaks = seq(0,6.0,1.5), limits = c(0,6.0), expand = c(0.1,0)) +
scale_fill_manual(
    values = c(
      "Sea star" = "#66c2a5",  # green
      "Crab" = "#fc8d62",      # orange
      "Both" = "#8da0cb"       # purple
    ),
    labels = c(
      "Sea star" = expression(italic(Pisaster)),
      "Crab" = "Rock crabs",
      "Both" = expression(italic(Pisaster) ~ "+ Rock crabs")
    )
  ) +
  labs(x = "Per capita interaction strength", y = NULL, fill = NULL) +
  theme_classic() +
  theme(
    legend.position = "top",  # Legend on top
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 20),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.text.y = element_text(size = 24, margin = margin(r = 20)),   
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.text.x = element_text(size = 20),
    axis.title.x = element_text(size = 24));PCIS

# PPIS plot with manual colors hardcoded
PPIS <- ggplot(ppis_all, aes(x = PPIS, y = Site, fill = Predator)) +
  geom_hline(yintercept = seq(1,3), linewidth = 1, color = "gray70") +
  geom_point(shape = 21, size = 7, position = position_dodge(0.0)) +
  scale_x_continuous(breaks = seq(0,1,0.2), limits = c(0,1), expand = c(0.1,0)) +
scale_fill_manual(
    values = c(
      "Sea star" = "#66c2a5",  # green
      "Crab" = "#fc8d62",      # orange
      "Both" = "#8da0cb"       # purple
    ),
    labels = c(
      "Sea star" = expression(italic(Pisaster)),
      "Crab" = "Rock crabs",
      "Both" = expression(italic(Pisaster) ~ "+ Rock crabs")
    )
  ) +
  labs(x = "Per population interaction strength", y = NULL, fill = NULL) +
  theme_classic() +
  theme(
    legend.position = "top",  # Legend on top
    legend.text = element_text(size = 18),
    legend.title = element_text(size = 20),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.text.y = element_text(size = 24, margin = margin(r = 20)),   
    axis.ticks.x = element_blank(),
    axis.line.x = element_blank(),
    axis.text.x = element_text(size = 20),
    axis.title.x = element_text(size = 24));PPIS


# Add tags to individual plots
PCIS <- PCIS + labs(tag = "A") +
  theme(
    plot.tag = element_text(size = 24)  # adjust tag style
  )

PPIS <- PPIS + labs(tag = "B") +
  theme(
    plot.tag = element_text(size = 24)
  )

# Extract shared legend from PCIS_plot
shared_legend <- get_legend(
  PCIS + theme(legend.position = "top", legend.text = element_text(size = 18), legend.title = element_blank())
)

# Remove legends from plots now that legend is shared
PCIS <- PCIS + theme(legend.position = "none")
PPIS <- PPIS + theme(legend.position = "none")

# Add vertical spacer
spacer <- ggdraw() + theme_void()

# Combine plots with shared legend on top
interactions_cow <- plot_grid(
  shared_legend,                             # legend at top
  plot_grid(
    PCIS,
    spacer,
    PPIS,
    ncol = 1,
    rel_heights = c(1, 0.2, 1)              # adjust spacer height
  ),
  ncol = 1,
  rel_heights = c(0.1, 1)                    # legend relative height
)

# Display combined figure
interactions_cow

```

