---
title: "MPE Manuscript Analysis 2021"
author: "Wesley Hull"
date: "11/7/2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r, echo=FALSE}
# Packages
library(tidyverse)
library(plyr)
library(Rmisc)
library(ggplot2)
library(RColorBrewer)
library(patchwork)
library(ggpubr)
library(glmmTMB)
library(DHARMa)
library(MuMIn)
library(lme4)
library(lsmeans)
library(car)
library(cowplot)
library(dplyr)
library(emmeans)
```

## Data Files Needed in Analysis 

  Recruitment data for R 2020  
  
  Field Density Data for R 2019
  
  Mussel.Survival.Data
  
  Pisaster.Measurements.All.2015
  
  Crab.Measurements.All.2015
  
```{r, echo=FALSE}
Recruitment_data_for_R_2020 <- read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Recruitment%20data%20for%20R%202020.csv")
Field_Density_Data_For_R_2019 <- read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Field%20Density%20Data%20For%20R%202019.csv")
Mussel_Survival_Data <- read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Mussel.Survival.Data.csv")
Pisaster_Measurements_All_2015 <-read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Pisaster.Measurements.All.2015.csv")
Crab_Meausrements_All_2015 <- read_csv("https://raw.githubusercontent.com/WesHull/MPE_Detection_Mode_Proj/refs/heads/main/Data/Crab.Meausrements.All.2015.csv")
```

# RECRUITMENT ANALYSIS 
```{r, echo = TRUE}
# New column dividing recruit count by days in the field
  Recruitment_data_for_R_2020$Rec.Tuff.Day<-Recruitment_data_for_R_2020$Count/104
  Recruitment_data_for_R_2020$Log.Rec.Day<-log(Recruitment_data_for_R_2020$Rec.Tuff.Day)
  
# Full model
  Rec.mod <- glmmTMB(Log.Rec.Day ~ Orientation * Site + (1 | Site:Plot), family = gaussian(), data = Recruitment_data_for_R_2020)
  summary(Rec.mod)
  Anova(Rec.mod)

    # interaction not significant
    # significant main effects of orientation and site
  
  # diagnostics
  rec.res <- simulateResiduals(fittedModel = Rec.mod)
  plot(rec.res)

# post-hoc
  # orientation
  emmeans(Rec.mod, pairwise ~ Orientation)
  
    # recruitment was higher on horizontal vs vertical surfaces
  
  # site
  emmeans(Rec.mod, pairwise ~ Site, adjust = "tukey")

    # Recruitment was highest at Point st. George, but did not differ between Devil's gate and Belinda point

```

# DENSITY ANALYSIS 
```{r, echo=TRUE}
# New column with log transformed Density data 
  Field_Density_Data_For_R_2019$Log.Density<-log(Field_Density_Data_For_R_2019$Density)
  head(Field_Density_Data_For_R_2019)
  
# Full Model
  Density.mod<-glmmTMB(Log.Density~Predator*Site, family = gaussian(), data = Field_Density_Data_For_R_2019)
  summary(Density.mod)
  Anova(Density.mod)

    # significant effect of predator and site
    
  # diagnostic
  plot(simulateResiduals(Density.mod))

# post-hoc
  # site
  emmeans(Density.mod, pairwise ~ Site, adjust = "tukey")
  
    # predator density was highest at PSG then DG then BP
  
  # predator
  emmeans(Density.mod, pairwise ~ Predator)

    # Sea star density is higher than crab density
```
  
# MUSSEL SURVIVAL ANALYSES 
### Proportion of Living 
```{r, echo=TRUE}
# Random affect is nested: blocks are repeated within sites, so code should be (1|Block:Site) or (1|Site:Block).
    
# With random intercept
    Mod1<-lmer(Proportion.Living~Treatment.S*Treatment.C*Site+(1|Site:Block), data = Mussel_Survival_Data)
    summary(Mod1)       #Variation from block is extremely low
    
# Without random intercept
    Mod2<-lm(Proportion.Living~Treatment.S*Treatment.C*Site,data = Mussel_Survival_Data)
    summary(Mod2)
      
# Likelihood ratio test
    anova(Mod1,Mod2)    # No difference between models, 
                        # Drop effect of block from model
    
# Model Selection
    
  # full
    Full.lm<-lm(Proportion.Living~Treatment.S*Treatment.C*Site, data = Mussel_Survival_Data)
      
# model comparison function
    options(na.action=na.fail)
    
    dredge(Full.lm)   # Full model is the best model
      
# Model summary/anova
    summary(Full.lm)
    anova(Full.lm)      # Run pairwise comparisons
    
# Pairwise comparisons
    lsmeans(Full.lm, pairwise ~ Site*Treatment.C*Treatment.S, adjust = "tukey")
  
```

### Natural Log Mussel Survival per Day 
```{r, echo=TRUE}
# Random effect assessment
# Random affect is nested: blocks are repeated within sites, so code should be (1|Block:Site) or (1|Site:Block).
      
# With random intercept
    Mod3<-lmer(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site+(1|Site:Block), data = Mussel_Survival_Data)
    summary(Mod3)       # Seems that the variation from block is extremely low
      
# Without random intercept
    Mod4<-lm(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site,data = Mussel_Survival_Data)
    summary(Mod4)
      
# Likelihood ratio test
    anova(Mod3,Mod4)    # No difference between models, so drop effect of block from model
      
# Model Selection
  
# full
    Full.LN.lm<-lm(LN.Rate.Survival.Day~Treatment.S*Treatment.C*Site, data = Mussel_Survival_Data)
      
# model selection
    options(na.action=na.fail)
      
    dredge(Full.LN.lm)   # Full model is the best model
      
# Model summary/anova
    summary(Full.LN.lm)
    anova(Full.LN.lm)  
    
# Pairwise comparisons
    lsmeans(Full.lm, pairwise ~ Site*Treatment.C*Treatment.S, adjust = "tukey")
```
  
# INTERACTION STRENGTH ANALYSIS 
### Per Population Interaction Strength 
#### How much does the entire predator population at a site reduce mussel survival?
```{r, echo=TRUE}
# Remove control treatments    
  Mussel.Survival<-Mussel_Survival_Data[-c(19,20,21,22,23,24,43,44,45,46,47,48,67,68,69,70,71,72),]

# Full Model
    Full.PerCap.glm<-glm(Mussel.Survival$Per.Cap.Int.Str~Mussel.Survival$Treatments.All*Mussel.Survival$Site, family = gaussian())
    summary(Full.PerCap.glm)
    
# ANOVA
    Full.PerCap.aov<-aov(Full.PerCap.glm)
    summary(Full.PerCap.aov)      # Interaction is significant
    
# Tukey's 
    TukeyHSD(Full.PerCap.aov)

```

### Per Capita Interaction Strength 
```{r, echo=TRUE}    
# Full Model
    Full.Pop.lm<-lm(Mussel.Survival$Pop.Int.Str~Mussel.Survival$Treatments.All*Mussel.Survival$Site)
    summary(Full.Pop.lm)
    
# ANOVA
    Full.Pop.aov<-aov(Full.Pop.lm)
    summary(Full.Pop.aov)      # Interaction is significant
    
# Tukey's 
    TukeyHSD(Full.Pop.aov) 
```
    
# SUPPLEMENTAL ANALYSIS 
```{r, echo=TRUE}
# ANOVA
# Sea star
  Rad<-lm(`Star Radius`~ Site, data = Pisaster_Measurements_All_2015)
  summary(Rad)
  Rad.aov<-aov(Rad)
  summary(Rad.aov)
  
# Crabs
  CP<-lm(`Carapace Width (mm)` ~ Site, data = Crab_Meausrements_All_2015)  
  summary(CP)
  CP.aov<-aov(CP)
  summary(CP.aov)  
  
  par(mfrow = c(2,2))
  plot(Rad.aov)   #violates assumptions, run kruskal wallis
  plot(CP.aov)    #violates assumptions, run kruskal wallis
    
# Kruskal wallis
  NP<-kruskal.test(Pisaster_Measurements_All_2015$Site ,Pisaster_Measurements_All_2015$`Star Radius`)
  NP  
  CP<-kruskal.test(Crab_Meausrements_All_2015$Site,Crab_Meausrements_All_2015$`Carapace Width (mm)`)
  CP
```  




# model selection, PCIS and PPIS
```{r}
# PPIS = total population level impact of each predator type per site
  # PPIS = PPIS=(effect of predator on survival(effect size))×(predator abundance)
    # effect size  =  survival in the respective absent treatment - the survival in the predator treatment

# 1. FULL FACTORIAL MODEL
  # convert predator treatment and site to factors
  Mussel_Survival_Data_No_Control <- Mussel_Survival_Data_No_Control %>%
    mutate(
      Site = factor(Site),
      Treatments.All = factor(Treatments.All),
      Block = factor(Block),
      Site.Block = interaction(Site, Block)
    )

  # binomial distribution
  mod_full <- glmmTMB(
    cbind(Survived, 30 - Survived) ~ Site * Treatments.All +
      (1 | Site:Block),
    family = binomial,
    data = Mussel_Survival_Data_No_Control
  )
  
  summary(mod_full)
  plot(simulateResiduals(mod_full))
  testDispersion(mod_full)
  
  # beta-binomial model
  mod_bb <- glmmTMB(
    cbind(Survived, 30 - Survived) ~
      Site * Treatments.All +
      (1 | Site:Block),
    family = betabinomial(link = "logit"),
    data = Mussel_Survival_Data_No_Control
  )
  
  Anova(mod_bb, type = "II")
    
    summary(mod_bb)
    plot(simulateResiduals(mod_bb))
  
    # Strong spatial heterogeneity (Site dominates system)
    # Crabs are the dominant predator effect
    # Sea stars are weak / inconsistent
    # No predator synergy
    # Crab effects vary by site (context dependence)
  
  # Post-hoc comparison
  # Treatment differences within each site
  pairs(emmeans(mod_bb, ~ Treatments.All | Site, type = "response"))
  
  # Site differences within each treatment
  pairs(emmeans(mod_bb, ~ Site | Treatments.All, type = "response"))
    
    
# 1. EMMs from new model
  emm <- emmeans(
    mod_bb,
    ~ Treatments.All | Site,
    type = "response"
  )
  
  emm_df <- as.data.frame(emm)

# 2. Extract treatment means + SE
  # have to add back in the control survival values
  ppis <- emm_df %>%
    select(Site, Treatments.All, prob, SE) %>%
    mutate(Group = case_when(
      Treatments.All == "Rock crabs"            ~ "crab",
      Treatments.All == "Pisaster"              ~ "star",
      Treatments.All == "Pisaster + Rock crabs" ~ "both"
    )) %>%
    select(Site, Group, prob, SE) %>%
    tidyr::pivot_wider(
      names_from = Group,
      values_from = c(prob, SE)
    ) %>%
    mutate(
      prob_control = 1.0,
      SE_control = 0.0
    )

# 3. Compute PPIS effects + SE propagation
  ppis <- ppis %>%
    mutate(
      PPIS_crab = prob_control - prob_crab,
      PPIS_star = prob_control - prob_star,
      PPIS_both = prob_control - prob_both,
      SE_crab = sqrt(SE_control^2 + SE_crab^2),
      SE_star = sqrt(SE_control^2 + SE_star^2),
      SE_both = sqrt(SE_control^2 + SE_both^2)
    )
  
# 4. Reshape to long format (SAFE + aligned)
  ppis_long <- ppis %>%
    select(
      Site,
      PPIS_crab, PPIS_star, PPIS_both,
      SE_crab, SE_star, SE_both
    ) %>%
    tidyr::pivot_longer(
      cols = starts_with("PPIS_"),
      names_to = "Predator",
      values_to = "PPIS"
    ) %>%
    mutate(
      SE = case_when(
        Predator == "PPIS_crab" ~ SE_crab,
        Predator == "PPIS_star" ~ SE_star,
        Predator == "PPIS_both" ~ SE_both
      ),
      Predator = dplyr::recode(
        Predator,
        PPIS_crab = "Crab",
        PPIS_star = "Sea star",
        PPIS_both = "Both"
      )
    ) %>%
    select(Site, Predator, PPIS, SE)
  
  ppis_long <- ppis_long %>%
    mutate(
      PPIS_lower = PPIS - 1.96 * SE,
      PPIS_upper = PPIS + 1.96 * SE
    )
  
# 5. Field predator densities (site means)
  density_site <- Field_Density_Data_For_R_2019 %>%
    dplyr::group_by(Site, Predator) %>%
    dplyr::summarise(
      density = mean(Density, na.rm = TRUE),
      .groups = "drop"
    )
  
  # combined predator density ("Both")
  density_both <- density_site %>%
    dplyr::group_by(Site) %>%
    dplyr::summarise(
      density = sum(density),
      .groups = "drop"
    ) %>%
    dplyr::mutate(Predator = "Both")
  
  density_all <- dplyr::bind_rows(density_site, density_both)
  
# 6. Merge PPIS + density

  ppis_long$Site <- factor(ppis_long$Site, ordered = FALSE)
  density_all$Site <- factor(density_all$Site, ordered = FALSE)
  
  ppis_long <- ppis_long %>%
    left_join(density_all, by = c("Site", "Predator"))
  
# 7. Compute PCIS + uncertainty
  ppis_long <- ppis_long %>%
    mutate(
      PCIS = PPIS / density,
      SE_PCIS = SE / density,
      PCIS_lower = PCIS - 1.96 * SE_PCIS,
      PCIS_upper = PCIS + 1.96 * SE_PCIS
    )
  
# 8. Final output
  ppis_long

```


##########################
# FIGURES
##########################

# Mussel recruitment
```{r}
Recruitment_data_for_R_2020$Rec.Tuff.Day<-Recruitment_data_for_R_2020$Count/104
Recruitment_data_for_R_2020$Log.Rec.Day<-log(Recruitment_data_for_R_2020$Rec.Tuff.Day)
head(Recruitment_data_for_R_2020)

# factorize site
Recruitment_data_for_R_2020$Site=factor(Recruitment_data_for_R_2020$Site, levels = c("Point St. George","Devil's Gate","Belinda Point"), ordered = TRUE)


# orientation
Rec.or<-ggplot(Recruitment_data_for_R_2020, aes(x=`Orientation`, y = `Rec.Tuff.Day`))+
  geom_boxplot(colour = "black", shape = 19)+
  scale_y_continuous(breaks=seq(0.0,9,3),limits = c(0.0,9), expand = c(0,0))+
  #scale_fill_manual(values = c("#CC79A7","#E69F00"))+
  labs(x="", fill ="")+
  ylab(expression("Mussels " %.% collector^{-1} %.% day^{-1}))+
  theme_classic()+
  theme(axis.ticks = element_line(linewidth = 0.5),legend.position = c(.83, .93),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 14))
Rec.or

Rec.site<-ggplot(Recruitment_data_for_R_2020, aes(x=`Site`, y = `Rec.Tuff.Day`))+
  geom_boxplot(colour = "black", shape = 19)+
  scale_y_continuous(breaks=seq(0.0,9,3),limits = c(0.0,9), expand = c(0,0))+
  #scale_fill_manual(values = c("#CC79A7","#E69F00"))+
  labs(x="", fill ="")+
  ylab(expression("Mussels " %.% collector^{-1} %.% day^{-1}))+
  theme_classic()+
  theme(axis.ticks = element_line(linewidth = 0.5),legend.position = c(.83, .93),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 14))
Rec.site

#ggsave("Rec.Fig.png", width = 12, height = 10, units = c("in"), dpi = 600)
```


# Figure Field Density
```{r}
###########
# averages
###########
Dens.Sum<-summarySE(Field_Density_Data_For_R_2019, measurevar = "Density", groupvars = c("Predator","Site"))
Dens.Sum$Site=factor(Dens.Sum$Site, levels = c("Point St. George","Devil's Gate","Belinda Point"), ordered = TRUE)
Dens.Sum$Predator=factor(Dens.Sum$Predator, levels = c("Sea star","Crab"), ordered = TRUE)

Den<-ggplot(Dens.Sum, aes(x=Site, y = Density, fill = Predator))+
  #geom_bar(stat = "summary",fun = "mean", color = "black", size = 1, position = position_dodge())+
  geom_errorbar(aes(ymin=Density-se, ymax=Density+se), width =0,size = 0.8, position = position_dodge(0.25), color = "grey40")+
  geom_point(size = 7, shape = 21, position = position_dodge(0.25))+
  scale_fill_manual(values = c("#CC79A7","#E69F00"), labels = c(expression(italic(Pisaster),"Rock crabs")))+
  labs(x="", fill ="Predators")+
  scale_y_continuous(expand = c(0,0), breaks=seq(0.0,0.3,0.1,),limits = c(0.0,0.3))+
  ylab(expression(paste("Predators " %.% m^-2, sep = "")))+
  theme_classic()+
  #border(size = 1.3)+
  theme(axis.ticks = element_line(size = 1.0), legend.position = c(.86,.91))+
  My_Theme

Den

#ggsave("Density.Fig.png", width = 12, height = 12, units = c("in"), dpi = 600)

###########
# boxplots
###########
Field_Density_Data_For_R_2019$Site=factor(Field_Density_Data_For_R_2019$Site, levels = c("Point St. George","Devil's Gate","Belinda Point"), ordered = TRUE)
Field_Density_Data_For_R_2019$Predator=factor(Field_Density_Data_For_R_2019$Predator, levels = c("Sea star","Crab"), ordered = TRUE)


Den.site <- ggplot(
  Field_Density_Data_For_R_2019,
  aes(x = Site, y = Density)) +
  geom_boxplot(
    width = 0.65,
    alpha = 0.8,
    outlier.shape = NA,
    position = position_dodge(width = 0.75),
    color = "black") +
  labs(x = "") +
  scale_x_discrete(labels = c(
  "Point St. George" = "PSG",
  "Devil's Gate" = "DG",
  "Belinda Point" = "BP"))+
  scale_y_continuous(
    expand = c(0, 0),
    breaks = seq(0.0, 0.3, 0.1),
    limits = c(0.0, 0.3)) +
  ylab(expression(paste("Predators " %.% m^-2, sep = ""))) +
  theme_classic() +
  theme(axis.ticks = element_line(linewidth = 0.5),legend.position = c(.83, .93),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 14));Den.site

Den.pred <- ggplot(
  Field_Density_Data_For_R_2019,
  aes(x = Predator, y = Density)) +
  geom_boxplot(
    width = 0.65,
    alpha = 0.8,
    outlier.shape = NA,
    position = position_dodge(width = 0.75),
    color = "black") +
  labs(x = "") +
  scale_y_continuous(
    expand = c(0, 0),
    breaks = seq(0.0, 0.3, 0.1),
    limits = c(0.0, 0.3)) +
  scale_x_discrete(labels = c(expression(italic("Pisaster"), "Rock crabs")))+
  ylab(expression(paste("Predators " %.% m^-2, sep = ""))) +
  theme_classic() +
  theme(axis.ticks = element_line(linewidth = 0.5),legend.position = c(.83, .93),
    axis.text.y = element_text(size = 12),
    axis.title.y = element_text(size = 14), 
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 14));Den.pred
```

# Merge recruitment and density
```{r}
# Add tags to individual plots
Rec.site <- Rec.site + labs(tag = "A") +
  theme(
    plot.tag = element_text(size = 14),
    axis.text.x = element_text(color = "white"),
    plot.margin = margin(b = 30)
  )

Rec.or <- Rec.or + labs(tag = "B") +
  theme(
    plot.tag = element_text(size = 14),
    axis.title.y = element_text(color = "white"),
    axis.text.y = element_text(color = "white"),
    plot.margin = margin(b = 30)
  )

Den.site <- Den.site + labs(tag = "C") +
  theme(
    plot.tag = element_text(size = 14)
  )

Den.pred <- Den.pred + labs(tag = "D") +
  theme(
    plot.tag = element_text(size = 14),
    axis.title.y = element_text(color = "white"),
    axis.text.y = element_text(color = "white"),
  )

# Add vertical spacer
spacer <- ggdraw() + theme_void()

# Combine plots with shared legend on top
RecDen_cow <- plot_grid(
  Rec.site,
  Rec.or,
  Den.site,
  Den.pred,
  ncol = 2,
  nrow = 2,
  align = "hv",
  axis = "tblr")

# Display combined figure
RecDen_cow
```

# Proportion Survival
```{r}
Field.Sum<-summarySE(Field_Consumption_Data_2019, measurevar = "Proportion.Living", groupvars = c("Site","Treatment.C","Treatment.S","Group"))
head(Field.Sum)

# Factorize site
Field.Sum$Site=factor(Field.Sum$Site, levels = c("Point St. George","Devil's Gate","Belinda Point"), ordered = TRUE)

#Subset data
PSG<-Field.Sum[9:12,]
DG<-Field.Sum[5:8,]
BP<-Field.Sum[1:4,]

# Point St. George figure
# Note: all mussels survive in control treatments, so there is no error bar for -crab, -star
PSG.Fig<-ggplot(data = PSG, 
               aes(x=`Treatment.S`, 
                   y=`Proportion.Living`, 
                   group = `Treatment.C`,
                   fill = `Treatment.C`))+
  geom_errorbar(aes(ymin=Proportion.Living-se, ymax=Proportion.Living+se),
                color = "grey40", 
                width =0,size = 0.8, 
                position = position_dodge(.9))+
geom_point(aes(fill = interaction(Treatment.C, Treatment.S)),
           size = 5,
           shape = 21,
           color = "black",
           stroke = 1,
           position = position_dodge(.9)) +
scale_fill_manual(values = c(
  "- Rock crabs.- Pisaster" = "white",
  "+ Rock crabs.- Pisaster" = "#E69F00",
  "- Rock crabs.+ Pisaster" = "#CC79A7",
  "+ Rock crabs.+ Pisaster" = "#009E73"
))+
scale_fill_manual(
  values = c(
    "- Rock crabs.- Pisaster" = "white",
    "+ Rock crabs.- Pisaster" = "#E69F00",
    "- Rock crabs.+ Pisaster" = "#CC79A7",
    "+ Rock crabs.+ Pisaster" = "#009E73"
  ),
  labels = c(
    "- Rock crabs.- Pisaster" = "Predators absent",
    "+ Rock crabs.- Pisaster" = "Rock crabs",
    "- Rock crabs.+ Pisaster" = expression(italic(Pisaster)),
    "+ Rock crabs.+ Pisaster" = expression(italic(Pisaster)~"+ Rock crabs")
  )
)+

scale_x_discrete(labels=expression("- "~italic(Pisaster),
                                     "+ "~italic(Pisaster)))+
  scale_y_continuous(breaks=seq(0.0,1.0,0.5,),
                     limits = c(0.0,1.0))+
  labs(x="",
       y="Proportion of live mussels", 
       group = "Predator", 
       fill = "")+
  theme_classic()+
  ggtitle("Point St. George")+
  theme(title = element_text(size = 24, color = "black"), 
        axis.tick.x = element_blank(), 
        legend.position = "right",
        legend.text = element_text(color = "black",size = 20),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5,size = 26),
        axis.text.x = element_blank(),
        axis.title.y = element_text(),
        axis.text.y = element_text(color = "black", size = 20),
        axis.title.y.left = element_text(color = "black",size = 24),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5))
  
PSG.Fig

#ggsave("PSG.Survival.Fig.png", width = 7.5, height = 7, units = c("in"), dpi = 600)


# Devils gate figure
# No mussels died in the control plots so no measure of error for -crab, -star treatments
DG.Fig<-ggplot(data = DG, 
               aes(x=`Treatment.S`, 
                   y=`Proportion.Living`, 
                   group = `Treatment.C`,
                   fill = `Treatment.C`))+
  geom_errorbar(aes(ymin=Proportion.Living-se, 
                    ymax=Proportion.Living+se),
                color = "grey40", 
                width =0,
                size = 0.8, 
                position = position_dodge(0.0))+
  geom_line(aes(color=Treatment.C), 
            size = 1.0, 
            color = "grey40")+
  geom_point(aes(color = Treatment.C), 
             size=5,
             shape=21,
             color = "black",
             stroke = 1)+
  scale_fill_manual(values = c("white","#E69F00"))+
  scale_x_discrete(labels=expression("- "~italic(Pisaster),
                                     "+ "~italic(Pisaster)))+
  scale_y_continuous(breaks=seq(0.0,1.0,0.5,),
                     limits = c(0.0,1.0))+
  labs(x="",
       y="", 
       group = "Predator")+
  theme_classic()+
  ggtitle("Devil's Gate")+
  theme(title = element_text(size = 24, color = "black"), 
        axis.ticks = element_line(size = 1.0), 
        legend.position = "bottom",
        legend.text = element_text(color = "black",size = 20),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5,size = 26),
        axis.text.x = element_text(color = "black", size = 20),
        axis.title.y = element_text(),
        axis.text.y = element_text(color = "black", size = 20),
        axis.title.y.left = element_text(color = "black",size = 24),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5))

DG.Fig

#ggsave("DG.Survival.Fig.png", width = 7.5, height = 7, units = c("in"), dpi = 600)


BP.Fig<-ggplot(data = BP, 
               aes(x=`Treatment.S`, 
                   y=`Proportion.Living`, 
                   group = `Treatment.C`,
                   fill = `Treatment.C`))+
  geom_errorbar(aes(ymin=Proportion.Living-se, 
                    ymax=Proportion.Living+se),
                color = "grey40", 
                width =0,
                size = 0.8, 
                position = position_dodge(0.0))+
  geom_line(aes(color=Treatment.C), 
            size = 1.0, color = "grey40")+
  geom_point(aes(color = Treatment.C), 
             size=5,
             shape=21,
             color = "black",
             stroke = 1)+
  scale_fill_manual(values = c("white","#E69F00"))+
  scale_x_discrete(labels=expression("- "~italic(Pisaster),
                                     "+ "~italic(Pisaster)))+
  scale_y_continuous(breaks=seq(0.0,1.0,0.5,),
                     limits = c(0.0,1.0))+
  labs(x="",
       y="", 
       group = "Predator")+
  theme_classic()+
  ggtitle("Belinda Point")+
  theme(title = element_text(size = 24, color = "black"), 
        axis.ticks = element_line(size = 1.0), 
        legend.position = "bottom",
        legend.text = element_text(color = "black",size = 20),
        legend.title = element_blank(),
        plot.title = element_text(hjust = 0.5,size = 26),
        axis.text.x = element_text(color = "black", size = 20),
        axis.title.y = element_text(),
        axis.text.y = element_text(color = "black", size = 20),
        axis.title.y.left = element_text(color = "black",size = 24),
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5))

BP.Fig

# Now arrange with common legend
pred.fig <-
ggarrange(PSG.Fig, DG.Fig, BP.Fig, common.legend = TRUE, ncol = 3, nrow = 1);pred.fig
















# Create interaction factor
PSG$Treatment_combo <- interaction(
  PSG$Treatment.C,
  PSG$Treatment.S,
  sep = " "
)

# FIXED color map (MATCHES YOUR REAL LEVELS)
cols <- c(
  "- Rock crabs - Pisaster" = "white",
  "- Rock crabs + Pisaster" = "#CC79A7",
  "+ Rock crabs - Pisaster" = "#E69F00",
  "+ Rock crabs + Pisaster" = "#009E73"
)

# Plot
PSG.Fig <- ggplot(PSG,
                   aes(x = Treatment.S,
                       y = Proportion.Living,
                       group = Treatment_combo)) +

  geom_errorbar(aes(ymin = Proportion.Living - se,
                    ymax = Proportion.Living + se),
                color = "grey40",
                width = 0,
                size = 0.8) +

  geom_line(aes(color = Treatment_combo),
            size = 1.0) +

  geom_point(aes(fill = Treatment_combo),
             shape = 21,
             size = 5,
             color = "black",
             stroke = 1) +

  scale_fill_manual(values = cols, name = "Treatment") +
  scale_color_manual(values = cols, name = "Treatment") +

  theme(legend.position = "right")
PSG.Fig




summary_data<-summarySE(Mussel_Survival_Data, measurevar = "Proportion.Living", groupvars = c("Site","Treatments.All"))
head(summary_data)

# new predator figure
pred.fig <-
ggplot(Mussel_Survival_Data, aes(x = Treatments.All, y = Proportion.Living, fill = Treatments.All))+
  geom_jitter(width = 0.2, height = 0, size = 2, alpha = 0.4, shape = 21)+
  geom_errorbar(data = summary_data,
                aes(ymin = Proportion.Living - se, ymax = Proportion.Living + se),
                width = 0, size = 0.8, color = "grey40")+
  geom_point(data = summary_data,
             aes(y = Proportion.Living),
             size = 4, shape = 21, color = "black", stroke = 1)+
  scale_y_continuous(breaks = seq(0,1,0.25), limits = c(0,1))+
  scale_x_discrete(limits = c("Control", "Pisaster", "Rock crabs", "Pisaster + Rock crabs"))+
  scale_fill_manual(
    values = c(
      "Control" = "white",
      "Rock crabs" = "#E69F00",
      "Pisaster" = "#CC79A7",
      "Pisaster + Rock crabs" = "#009E73"),
    breaks = c("Control", "Pisaster", "Rock crabs", "Pisaster + Rock crabs"),
    labels = c("Control",
               expression(italic("Pisaster")),
               "Rock crabs",
               expression(italic("Pisaster") ~ "+ Rock crabs")))+
  labs(x = "", y = "Proportion of living mussels", fill = "")+
  facet_wrap(~factor(Site, levels = c("Point St. George", "Devil's Gate", "Belinda Point")))+  
  theme(
    legend.position = "bottom",
    legend.text = element_text(size = 12),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.text.y  = element_text(size = 12, color = "black"),
    axis.title.y = element_text(size = 14),
    axis.line.y = element_line(),
    panel.background = element_blank(),
    panel.grid = element_blank(),
    strip.background = element_blank(),
    strip.text = element_text(size = 14)); pred.fig
```

# Figures: PPIS and PCIS
```{r}

# Recode Predator factor
ppis_long <- ppis_long %>%
  mutate(
    Predator = factor(Predator, levels = c("Both", "Crab", "Sea star"))
  )

# PCIS plot with manual colors hardcoded
PCIS<- ggplot(ppis_long, aes(x = PCIS, y = Site, fill = Predator)) +
  #geom_hline(yintercept = seq(1,3), linewidth = 1, color = "gray70") +
  geom_errorbarh(aes(xmin = PCIS_lower, xmax = PCIS_upper),height = 0,linewidth = .5, position = position_dodge(0.5), color = "grey40") +
  geom_point(shape = 21, size = 4, position = position_dodge(0.5)) +
  scale_x_continuous(breaks = seq(0,6.0,1.5), limits = c(0,6.0), expand = c(0.1,0)) +
scale_fill_manual(
    values = c(
      "Sea star" = "#CC79A7",  # green
      "Crab" = "#E69F00",      # orange
      "Both" = "#009E73"       # purple
    ),
    labels = c(
      "Sea star" = expression(italic(Pisaster)),
      "Crab" = "Rock crabs",
      "Both" = expression(italic(Pisaster) ~ "+ Rock crabs")
    )
  ) +
  labs(x = "Per capita interaction strength", y = NULL, fill = NULL) +
  theme_classic() +
  theme(
    legend.position = "top",  # Legend on top
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.text.y = element_text(size = 14, margin = margin(r = 20)),
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 14));PCIS

# PPIS plot with manual colors hardcoded
PPIS <- ggplot(ppis_long, aes(x = PPIS, y = Site, fill = Predator)) +
 # geom_hline(yintercept = seq(1,3), linewidth = 1, color = "gray70") +
  geom_errorbarh(aes(xmin = PPIS_lower, xmax = PPIS_upper),height = 0,linewidth = .5, position = position_dodge(0.5), color = "grey40") +
  geom_point(shape = 21, size = 4, position = position_dodge(0.5)) +
  scale_x_continuous(breaks = seq(0,1,0.2), limits = c(0,1), expand = c(0.1,0)) +
scale_fill_manual(
    values = c(
      "Sea star" = "#CC79A7",  # green
      "Crab" = "#E69F00",      # orange
      "Both" = "#009E73"       # purple
    ),
    labels = c(
      "Sea star" = expression(italic(Pisaster)),
      "Crab" = "Rock crabs",
      "Both" = expression(italic(Pisaster) ~ "+ Rock crabs")
    )
  ) +
  labs(x = "Per population interaction strength", y = NULL, fill = NULL) +
  theme_classic() +
  theme(
    legend.position = "top",  # Legend on top
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    axis.ticks.y = element_blank(),
    axis.line.y = element_blank(),
    axis.text.y = element_text(size = 14, margin = margin(r = 20)),
    axis.text.x = element_text(size = 12),
    axis.title.x = element_text(size = 14));PPIS


# Add tags to individual plots
PCIS <- PCIS + labs(tag = "B") +
  theme(
    plot.tag = element_text(size = 14)  # adjust tag style
  )

PPIS <- PPIS + labs(tag = "A") +
  theme(
    plot.tag = element_text(size = 14)
  )

# Extract shared legend from PCIS_plot
shared_legend <- get_legend(
  PCIS + theme(legend.position = "top", legend.text = element_text(size = 14), legend.title = element_blank())
)

# Remove legends from plots now that legend is shared
PCIS <- PCIS + theme(legend.position = "none")
PPIS <- PPIS + theme(legend.position = "none")

# Add vertical spacer
spacer <- ggdraw() + theme_void()

# Combine plots with shared legend on top
interactions_cow <- plot_grid(
  shared_legend,                             # legend at top
  plot_grid(
    PPIS,
    spacer,
    PCIS,
    ncol = 1,
    rel_heights = c(1, 0.2, 1)              # adjust spacer height
  ),
  ncol = 1,
  rel_heights = c(0.1, 1)                    # legend relative height
)

# Display combined figure
interactions_cow

```

# Final figures
```{r}
RecDen_cow
ggsave("Outputs/RecDen_cow.tiff", width = 7, height = 7, units = c("in"), dpi = 300)

pred.fig
ggsave("Outputs/pred.fig.tiff", width = 6, height = 6, units = c("in"), dpi = 300)

interactions_cow
ggsave("Outputs/interactions_cow.tiff", width = 6, height = 6, units = c("in"), dpi = 300)

```